---
title: "Effect of Lowering Activation Time"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r import libraries, include=TRUE}

library(ggplot2)
library(scales)
library(dplyr)    
library(resample)


```

The trial may be modeled as a doubly stochastic Poisson process (Cox process), in which the intensity function is according to the site activation times and recruitment rates for open sites. The trial may be simulated by first drawing activation times and recruitment rates, then drawing the time of the Nth recruitment as a time transformation of a rate-1 Poisson process.

```{r define function for simulating trial completion}


NSPP_Intensity <- function(t_val, activation_times, recruitment_rates){
  return(
    
    sum((t_val-activation_times) * recruitment_rates * (activation_times < t_val))
    
  )
}

NSPP_Intensity_dif <- function(t_val_1, t_val_2, activation_times, recruitment_rates){
  return(abs(t_val_2 - NSPP_Intensity(t_val_1, activation_times, recruitment_rates)))
}

Cox_Inverse_Rate <- function(t_gamma, I, lambda_I, alpha, beta, max_t){
  
  activation_times <- rexp(n = I , rate=lambda_I)
  recruitment_rates <- rgamma(n = I , shape=alpha, rate=beta)
  
  
  inverse_rate_result <- optimize(f=NSPP_Intensity_dif
                                  ,t_val_2 = t_gamma
                                  ,activation_times=activation_times
                                  ,recruitment_rates=recruitment_rates
                                  ,interval=c(0,max_t)
                                  ,maximum = FALSE)
  
  return(inverse_rate_result$minimum)
  
}

simulate_site_activation_Cox <- function(N, I, lambda_I, alpha, beta, max_t, n_sims=1){
  
  gamma_draws <- rgamma(n = n_sims, shape = N, rate = 1)
  
  time_draws <- sapply(gamma_draws,Cox_Inverse_Rate, I=I, lambda_I=lambda_I, alpha=alpha, beta=beta, max_t=max_t)
  
  
  sim_results <- data.frame(matrix(unlist(c(
                                      rep(I,n_sims),
                                      time_draws,
                                      rep(N,n_sims))), ncol=3, byrow=FALSE))
  colnames(sim_results) <- c("Sites_Initiated"
                             , "Day_Completed"
                             , "Required_Recruitment")
  
  return(sim_results)

}

```


```{r function to simulate recruitment and get delay costs}

simulate_site_activation_manual <- function(N, I, lambda_I, alpha, beta, n_sims=1){
    
  activation_times <- rexp(n = I * n_sims, rate=lambda_I)
  recruitment_rates <- rgamma(n = I * n_sims, shape=alpha, rate=beta)
  recruited <- data.frame(matrix(unlist(rep(0,I*n_sims)), ncol=n_sims, byrow=FALSE))
  recruit_sums <- colSums(recruited)
  
  completed <- recruit_sums >= N
  completed_prev <- rep(FALSE, n_sims)
  day_completed <- rep(0,n_sims)
  
  day <- 0
  
  while(sum(completed) < n_sims){
    
    activated <- activation_times < day
    site_stopped <- rep(completed, each=I)
    
    recruitments_day <- activated * (! site_stopped) * rpois(n=I * n_sims, lambda = recruitment_rates)
    recruitments_day_df <- data.frame(matrix(unlist(recruitments_day), ncol=n_sims, byrow=FALSE))
    recruited <- recruited + recruitments_day_df
    recruit_sums <- colSums(recruited)
    completed <- recruit_sums >= N
  
    day <- day + 1  
    
    for(i in 1:n_sims){
      if (completed[i] && ! completed_prev[i]){
        day_completed[i] = day
        completed_prev[i] = TRUE
      }
    }
    
  }
  
  activated_df <- data.frame(matrix(unlist(activated), ncol=n_sims, byrow=FALSE))
  
  sites_activated <- colSums(activated_df)
  
  
  sim_results <- data.frame(matrix(unlist(c(
                                      rep(I,n_sims),
                                      sites_activated,
                                      day_completed,
                                      rep(N,n_sims))), ncol=4, byrow=FALSE))
  colnames(sim_results) <- c("Sites_Initiated","Sites_Activated",
                             "Day_Completed","Required_Recruitment")

  return(sim_results)  
}

```


```{r verify Cox process}

alpha_level <- 2
beta_level <- 4

N_level <- 3000
T_d_level <- 20 * 7

n_sims_level <- 1000

one_site_activation_time = 4*7/5

cohort_size_level <- 10

mean_activation_time = one_site_activation_time * cohort_size_level

sim_results_manual <- simulate_site_activation_manual(N=N_level
                                                      , I=cohort_size_level
                                                      , lambda_I = 1/mean_activation_time
                                                      , alpha=alpha_level
                                                      , beta=beta_level
                                                      , n_sims=n_sims_level
                                                      )

colMeans(sim_results_manual)


sim_results_Cox <- simulate_site_activation_Cox(N=N_level
                                                  , I=cohort_size_level
                                                  , lambda_I = 1/mean_activation_time
                                                  , alpha=alpha_level
                                                  , beta=beta_level
                                                  , n_sims=n_sims_level
                                                  , max_t = 2400
                                                  )

colMeans(sim_results_Cox)


```


```{r}

dens_manual=density(sim_results_manual$Day_Completed)
# Plot y-values scaled by number of observations against x values
plot(dens_manual$x,length(sim_results_manual$Day_Completed)*dens_manual$y,type="l",xlab="Value",ylab="Count estimate")


dens_Cox=density(sim_results_Cox$Day_Completed)
# Plot y-values scaled by number of observations against x values
plot(dens_Cox$x,length(sim_results_Cox$Day_Completed)*dens_Cox$y,type="l",xlab="Value",ylab="Count estimate")

```

Here we run a simulation using the Cox process to determine the optimal number of sites to open
depending on activation and delay costs, to reproduce main results from the more manual simulation.

```{r Run Cox process simulation to find optimal site cohort}

results_df <- data.frame()
result_cols <- c("Length_of_Trial_in_Weeks",
                 "Needed_Recruitment",
                 "Activation_Cost_per_Site",
                 "Delay_Cost_per_Week",
                 "Cohort_Size",
                 "Alpha",
                 "Beta",
                 "Mean_Activation_Time_in_Weeks",
                 "Average_Day_Completed",
                 "Site_Activation_Cost",
                 "Average_Delay_Cost",
                 "Variance_Delay_Cost",
                 "Average_Total_Cost",
                 "Variance_Total_Cost",
                 "Number_Simulations")

activation_cost_levels <- c(1000, 5000, 10000)
delay_cost_levels <- c(50000 / 7,100000 / 7,150000 / 7)
cohort_size_levels <- c(1:20) * 10

alpha_level <- 2
beta_level <- 4

N_level <- 3000
T_d_level <- 20 * 7

n_sims_level <- 10000

one_site_activation_time = 4*7/5

ptm <- proc.time()

for(cohort_size_level in cohort_size_levels){
      
  mean_activation_time = one_site_activation_time * cohort_size_level
  
  sim_results <- simulate_site_activation_Cox(N=N_level
                                                  , I=cohort_size_level
                                                  , lambda_I = 1/mean_activation_time
                                                  , alpha=alpha_level
                                                  , beta=beta_level
                                                  , n_sims=n_sims_level
                                                  , max_t = 2400
                                                  )

  

  for(activation_cost_level in activation_cost_levels){
  
    for(delay_cost_level in delay_cost_levels){
      
      sim_results["Activation_Cost"] <- cohort_size_level * activation_cost_level
      
      sim_results["Delay_Cost"] <- ifelse(sim_results[["Day_Completed"]] > T_d_level,
                                          sim_results[["Day_Completed"]] - T_d_level,
                                          0) * delay_cost_level
      
      sim_results["Total_Cost"] <- sim_results["Activation_Cost"] + sim_results["Delay_Cost"]
      
      sim_means <- colMeans(sim_results)
      sim_vars <- colVars(sim_results)
    
      results_newline <- data.frame(matrix(ncol=length(result_cols),nrow=1))
      colnames(results_newline) <- result_cols
    
      results_newline[["Length_of_Trial_in_Weeks"]] <- T_d_level / 7
      results_newline[["Needed_Recruitment"]] <- N_level
      results_newline[["Activation_Cost_per_Site"]] <- activation_cost_level
      results_newline[["Delay_Cost_per_Week"]] <- delay_cost_level * 7
      results_newline[["Alpha"]] <- alpha_level
      results_newline[["Beta"]] <- beta_level
      results_newline[["Mean_Activation_Time_in_Weeks"]] <- mean_activation_time / 7
      results_newline[["Cohort_Size"]] <- cohort_size_level
      results_newline[["Average_Day_Completed"]] <- sim_means["Day_Completed"]
      results_newline[["Site_Activation_Cost"]] <- sim_means["Activation_Cost"]
      results_newline[["Average_Delay_Cost"]] <- sim_means["Delay_Cost"]
      results_newline[["Variance_Delay_Cost"]] <- sim_vars["Delay_Cost"]
      results_newline[["Average_Total_Cost"]] <- sim_means["Total_Cost"]
      results_newline[["Variance_Total_Cost"]] <- sim_vars["Total_Cost"]
      results_newline[["Number_Simulations"]] <- n_sims_level
      
      results_df <- rbind(results_df,results_newline)
      
      
    }
  }
  
  write.csv(results_df, file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_16.csv",row.names=FALSE)
  
  print("Cohort Size:")
  print(cohort_size_level)

}

proc.time() - ptm

```


```{r plot results of delay cost simulation}

results_df <- read.csv(file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_16.csv")


results_df_delay <- cbind("Delay",
                          results_df[c("Cohort_Size",
                                       "Activation_Cost_per_Site",
                                       "Delay_Cost_per_Week",
                                       "Average_Delay_Cost")])
colnames(results_df_delay) <- c("Cost_Type",
                                "Cohort_Size",
                                "Activation_Cost_per_Site",
                                "Delay_Cost_per_Week",
                                "Average_Cost")


results_df_activation <- cbind("Activation",
                          results_df[c("Cohort_Size",
                                       "Activation_Cost_per_Site",
                                       "Delay_Cost_per_Week",
                                       "Site_Activation_Cost")])
colnames(results_df_activation) <- c("Cost_Type",
                                "Cohort_Size",
                                "Activation_Cost_per_Site",
                                "Delay_Cost_per_Week",
                                "Average_Cost")

results_df_total <- cbind("Total",
                          results_df[c("Cohort_Size",
                                       "Activation_Cost_per_Site",
                                       "Delay_Cost_per_Week",
                                       "Average_Total_Cost")])
colnames(results_df_total) <- c("Cost_Type",
                                "Cohort_Size",
                                "Activation_Cost_per_Site",
                                "Delay_Cost_per_Week",
                                "Average_Cost")

results_df_cost_type <- rbind(results_df_delay,
                              results_df_activation,
                              results_df_total)

activation_cost_labels <- paste(dollar_format()(activation_cost_levels),"per Site")
names(activation_cost_labels) <- activation_cost_levels

delay_cost_labels <- paste(dollar_format()(delay_cost_levels * 7),"per Week")
names(delay_cost_labels) <- delay_cost_levels * 7

ggplot() +
  geom_line(data=results_df_cost_type, aes(x=Cohort_Size,y=Average_Cost, group=Cost_Type, color=Cost_Type)) +
  facet_grid(Activation_Cost_per_Site ~ Delay_Cost_per_Week,
             labeller=labeller(Activation_Cost_per_Site=activation_cost_labels,
                               Delay_Cost_per_Week=delay_cost_labels)) +
  xlab("Initiated Sites") +
  ylab("Average Cost") +
  scale_y_continuous(labels=scales::dollar_format())

ggsave("C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_recruitment_study_16_a.png")
  
```


```{r get optimal costs}

results_df <- read.csv(file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_16.csv")

optimal_sets <- results_df %>% 
    group_by(Length_of_Trial_in_Weeks,
             Needed_Recruitment,
             Alpha,
             Beta,
             Activation_Cost_per_Site,
             Delay_Cost_per_Week) %>% 
    slice(which.min(Average_Total_Cost)) %>%
    rename(Optimal_Cohort_Size=Cohort_Size) %>%
    select(c("Length_of_Trial_in_Weeks","Needed_Recruitment",
             "Alpha","Beta","Activation_Cost_per_Site",
             "Delay_Cost_per_Week","Optimal_Cohort_Size",
             "Average_Day_Completed","Average_Delay_Cost",
             "Variance_Delay_Cost","Site_Activation_Cost",
             "Average_Total_Cost","Variance_Total_Cost"))

optimal_sets

write.csv(optimal_sets, file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_16_opt.csv")

```


Here we introduce variation in the average activation time parameters, and demonstrate the effect of this
changing parameter on the optimal cohort size and cost.


```{r Cox process simulation with changing expected activation time}

results_df <- data.frame()
result_cols <- c("Length_of_Trial_in_Weeks",
                 "Needed_Recruitment",
                 "Activation_Cost_per_Site",
                 "Delay_Cost_per_Week",
                 "Cohort_Size",
                 "Alpha",
                 "Beta",
                 "One_Site_Activation_Time_in_Weeks",
                 "Mean_Activation_Time_in_Weeks",
                 "Average_Day_Completed",
                 "Site_Activation_Cost",
                 "Average_Delay_Cost",
                 "Variance_Delay_Cost",
                 "Average_Total_Cost",
                 "Variance_Total_Cost",
                 "Number_Simulations")

activation_cost_levels <- c(1000, 5000, 10000)
delay_cost_levels <- c(50000 / 7,100000 / 7,150000 / 7)
cohort_size_levels <- c(1:20) * 10
one_site_activation_time_levels <- c(1:(16*2))*7/5/2

alpha_level <- 2
beta_level <- 4

N_level <- 3000
T_d_level <- 20 * 7

n_sims_level <- 10000


ptm <- proc.time()


for(one_site_activation_time in one_site_activation_time_levels){

  print("One Site Activation Time:")
  print(one_site_activation_time / 7 * 5)
    
  for(cohort_size_level in cohort_size_levels){
      
    mean_activation_time = one_site_activation_time * cohort_size_level
        
    sim_results <- simulate_site_activation_Cox(N=N_level
                                                    , I=cohort_size_level
                                                    , lambda_I = 1/mean_activation_time
                                                    , alpha=alpha_level
                                                    , beta=beta_level
                                                    , n_sims=n_sims_level
                                                    , max_t = 2400
                                                    )
  
    for(activation_cost_level in activation_cost_levels){
    
      for(delay_cost_level in delay_cost_levels){
        
        sim_results["Activation_Cost"] <- cohort_size_level * activation_cost_level
        
        sim_results["Delay_Cost"] <- ifelse(sim_results[["Day_Completed"]] > T_d_level,
                                            sim_results[["Day_Completed"]] - T_d_level,
                                            0) * delay_cost_level
        
        sim_results["Total_Cost"] <- sim_results["Activation_Cost"] + sim_results["Delay_Cost"]
        
        sim_means <- colMeans(sim_results)
        sim_vars <- colVars(sim_results)
      
        results_newline <- data.frame(matrix(ncol=length(result_cols),nrow=1))
        colnames(results_newline) <- result_cols
      
        results_newline[["Length_of_Trial_in_Weeks"]] <- T_d_level / 7
        results_newline[["Needed_Recruitment"]] <- N_level
        results_newline[["Activation_Cost_per_Site"]] <- activation_cost_level
        results_newline[["Delay_Cost_per_Week"]] <- delay_cost_level * 7
        results_newline[["Alpha"]] <- alpha_level
        results_newline[["Beta"]] <- beta_level
        results_newline[["One_Site_Activation_Time_in_Weeks"]] <- one_site_activation_time / 7
        results_newline[["Mean_Activation_Time_in_Weeks"]] <- mean_activation_time / 7
        results_newline[["Cohort_Size"]] <- cohort_size_level
        results_newline[["Average_Day_Completed"]] <- sim_means["Day_Completed"]
        results_newline[["Site_Activation_Cost"]] <- sim_means["Activation_Cost"]
        results_newline[["Average_Delay_Cost"]] <- sim_means["Delay_Cost"]
        results_newline[["Variance_Delay_Cost"]] <- sim_vars["Delay_Cost"]
        results_newline[["Average_Total_Cost"]] <- sim_means["Total_Cost"]
        results_newline[["Variance_Total_Cost"]] <- sim_vars["Total_Cost"]
        results_newline[["Number_Simulations"]] <- n_sims_level
        
        results_df <- rbind(results_df,results_newline)
        
        
      }
    }
  
  write.csv(results_df, file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_17.csv",row.names=FALSE)

  #print("Cohort Size:")
  #print(cohort_size_level)

  }

    
}



proc.time() - ptm

```



```{r get optimal costs}

results_df <- read.csv(file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_17.csv")

optimal_sets <- results_df %>% 
    group_by(Length_of_Trial_in_Weeks,
             Needed_Recruitment,
             Alpha,
             Beta,
             One_Site_Activation_Time_in_Weeks,
             Activation_Cost_per_Site,
             Delay_Cost_per_Week) %>% 
    slice(which.min(Average_Total_Cost)) %>%
    rename(Optimal_Cohort_Size=Cohort_Size) %>%
    select(c("Length_of_Trial_in_Weeks",
             "Needed_Recruitment",
             "Alpha","Beta",
             "One_Site_Activation_Time_in_Weeks",
             "Activation_Cost_per_Site",
             "Delay_Cost_per_Week",
             "Optimal_Cohort_Size",
             "Average_Day_Completed",
             "Average_Delay_Cost",
             "Variance_Delay_Cost",
             "Site_Activation_Cost",
             "Average_Total_Cost",
             "Variance_Total_Cost"))

optimal_sets

write.csv(optimal_sets, file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_17_opt.csv")

```


Verifying results are consistent with previous analyses with constant expected activation time

```{r}

optimal_sets_1 <- read.csv(file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_16_opt.csv")

optimal_sets_2 <- read.csv(file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_17_opt.csv")

optimal_sets_1 %>%
  select("Activation_Cost_per_Site",
         "Delay_Cost_per_Week",
         "Optimal_Cohort_Size",
         "Average_Total_Cost")

optimal_sets_2  %>% 
  filter(One_Site_Activation_Time_in_Weeks == 4/5) %>% 
  select("Activation_Cost_per_Site",
         "Delay_Cost_per_Week",
         "Optimal_Cohort_Size",
         "Average_Total_Cost")

```




```{r plot optimal cohort size with changing expected activation time}

optimal_df <- read.csv(file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_17_opt.csv")



activation_cost_labels <- paste(dollar_format()(activation_cost_levels),"per Site")
names(activation_cost_labels) <- activation_cost_levels

delay_cost_labels <- paste(dollar_format()(delay_cost_levels * 7),"per Week")
names(delay_cost_labels) <- delay_cost_levels * 7

ggplot() +
  geom_line(data=optimal_df, aes(x=One_Site_Activation_Time_in_Weeks * 5,
                                 y=Optimal_Cohort_Size)) +
  facet_grid(Activation_Cost_per_Site ~ Delay_Cost_per_Week,
             labeller=labeller(Activation_Cost_per_Site=activation_cost_labels,
                               Delay_Cost_per_Week=delay_cost_labels)) +
  xlab("Expected Activation Time (Weeks per Five Sites)") +
  ylab("Optimal Cohort Size")

ggsave("C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_recruitment_study_17_a.png")
  
```



```{r plot optimal cost with changing expected activation time}

optimal_df <- read.csv(file = "C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_reruitment_study_17_opt.csv")



activation_cost_labels <- paste(dollar_format()(activation_cost_levels),"per Site")
names(activation_cost_labels) <- activation_cost_levels

delay_cost_labels <- paste(dollar_format()(delay_cost_levels * 7),"per Week")
names(delay_cost_labels) <- delay_cost_levels * 7

ggplot() +
  geom_line(data=optimal_df, aes(x=One_Site_Activation_Time_in_Weeks * 5,
                                 y=Average_Total_Cost)) +
  facet_grid(Activation_Cost_per_Site ~ Delay_Cost_per_Week,
             labeller=labeller(Activation_Cost_per_Site=activation_cost_labels,
                               Delay_Cost_per_Week=delay_cost_labels)) +
  xlab("Expected Activation Time (Weeks per Five Sites)") +
  ylab("Optimal Average Cost") +
  scale_y_continuous(labels=scales::dollar_format())

ggsave("C:/Users/mclef/Documents/Research/Clinical Trials/sim results/Cohort Recruitment Distribution/Exponential Establishment/cohort_recruitment_study_17_b.png")
  
```